\textcolor{red}{Simple} connectome inference using partial correlation from imaging calcium signal

### 2014-06-16 article.tex

\subsection{Causal discovery methodology}
% Introduce a custom solution to make the Y matrix asymmetric
% Introduce directivity
Since partial correlation is a symmetric measure, the causal mechanism behind the
interaction ($A$ directly causes $B$, $B$ directly causes $A$ or both) can not
be directly inferred\footnote{Also note that two other causal mechanisms might be
implied while having a non-zero partial coefficient: (i) a pair of variables
is induced by a common (hidden) variable; (ii) $A$ (respectively $B$) is
conditionally correlated to a hidden variable affecting $B$ (resp. $A$)
\cite{de2004discovery}.}.

By ignoring the directivity, we systematically make a true positive and a
false positive for direct link.  Hoping to orientate
correctly the most obvious directed links, we attempt to retrieve some causal
information, i.e. the directivity of the links, by computing and stacking with
our symmetric adjacency matrix a matrix of pairwise activation heuristics.
This method computes an anti-symmetric activation score $Z$ based the
variation of fluorescence signal of a neuron $j$ due to a neuron $i$ over all
time steps:
\[
\hat{Z}: z_{ij} = \sum_{t=1}^{T - 1}
    \mathbb{1}(x_{t+1,j} - x_{t, i} \in \left[f1, f2 \right]) -
    \mathbb{1}(x_{t+1,i} - x_{t, j} \in \left[f1, f2 \right])
\]
where $f_1$ and $f_2$ are parameters of the method.

>>>>

% Introduce a custom solution to make the Y matrix asymmetric
% Introduce directivity
Since partial correlation is a symmetric measure, the causal mechanism behind the
interaction ($A$ directly causes $B$, $B$ directly causes $A$ or both) can not
be directly inferred\footnote{Also note that two other causal mechanisms might be
implied while having a non-zero partial coefficient: (i) a pair of variables
is induced by a common (hidden) variable; (ii) $A$ (respectively $B$) is
conditionally correlated to a hidden variable affecting $B$ (resp. $A$)
\cite{de2004discovery}.}.
In order to retrieve causal information, we developed an
heuristics based on the following observation. The rise of fluorescence
indicates the activation of a neuron $i$. If a neuron $j$ have also
an increased of fluorescence after a slight delay, this could have been
caused by the neuron $i$. Thus, we count whenever
such events occurred between consecutive time steps
\[
\hat{Z}: z_{ij} = \sum_{t=1}^{T - 1}
    \mathbb{1}(x_j^{t+1} - x_i^t \in \left[f_1, f_2 \right]) -
    \mathbb{1}(x_i^{t+1} - x_j^t \in \left[f_1, f_2 \right]), \forall i, j \in V
\]
where $f_1$ and $f_2$ are parameters of the method.
Finally, the weighted average of correlation matrices is combined with $Z$ such
that $Z$ has a contribution of $0.3\%$ to the prediction of an edge.
